%%% ===============================================================================
\chapter{Application}\label{sec:application}
%%% ===============================================================================
\section{NFDI4ING Model Validation}\label{sec:application:nfdi4ing-model-validation}

To evaluate the practical applicability of the implemented extension, the workflow was applied to data structures aligned with the NFDI4ING metadata model. The goal was to validate whether schema-driven instance data authored in MetaConfigurator can be transformed into semantically consistent \gls{jsonld}/\gls{rdf} representations that follow the expected ontology terms and relationship patterns.

The validation workflow followed four steps: (1) authoring and refining JSON instance data and JSON Schema in MetaConfigurator, (2) generating and reviewing \gls{rml} mappings (including \gls{ai}-assisted drafts), (3) transforming source JSON into \gls{jsonld}, and (4) checking the resulting triples through the RDF authoring panel and \gls{sparql} queries. This procedure ensured that both structural correctness (schema compliance) and semantic correctness (expected predicates/classes and links) were inspected before accepting results.

In the tested scenarios, the integrated workflow enabled fast iteration between data modeling and semantic validation. Typical issues observed in early drafts included incomplete context declarations, incorrect unit/resource IRIs, and overly broad AI-generated mappings. These were resolved through user-guided refinement in the mapping editor and immediate feedback from the triples view and query dialog. The results indicate that the extension is suitable for NFDI4ING-oriented modeling tasks where users need to move from structured JSON authoring to ontology-aligned graph data within a single tool.
